A Critical Review of Sophistication and Algorithmic Model Selection

### 2015-01-05 Facticity/Facticity/paper.tex
....

\newcommand{\M}{\mathscr M}
\newcommand{\C}{\mathscr C}
\newcommand{\T}{\mathscr T}
\newcommand{\F}{\mathscr F}
\renewcommand{\P}{\mathscr P}
\newcommand{\K}{\mathscr K}
\newcommand{\X}{\mathscr X}
\newcommand{\B}{\mathbb B}
\newcommand{\D}{\Delta}
\newcommand{\N}{\mathbb N}
\newcommand{\tn}[1]{\textnormal{#1}}
\newcommand{\pair}[1]{\left\langle{#1}\right\rangle}
\newcommand{\concat}{\oplus}
\newcommand{\symb}[1]{\texttt{#1}}
\newcommand{\br}[1]{\overline{#1}}
\newcommand{\s}{\tn{soph}}

@@

\begin{abstract}
Kolmogorov complexity is a sound expression of the information content of a binary string, but it clashes with our intuitive notions of information: the strings with maximal information content are not the ones we would regard as interesting, or meaningful. One attempt to capture this idea of `interesting' information is \emph{sophistication}. It is a simple idea with many attempted formalizations: describe the string efficiently with a two-part coding approach, and take the complexity of the model as a measure of the amount of interesting information. Kolmogorov complexity is invariant to the ad-hoc choices in its construction: change them, and the result stays the same, up to a constant. We review the history of sophistication critically, asking whether invariance exists for the proposed sophistication functions as well. In most cases, we can show that it doesn't. We conjecture that no formalization of sophistication can be invariant, and we briefly review the consequences of this conjecture.
\end{abstract}

@@

\noindent Kolmogorov complexity is an objective measure of the information content of an object. Although its definition contains several arbitrary choices, these only affect its value up to a constant additive term. This property is crucial for the measure's success: it allows us to view Kolmogorov complexity as a \emph{property of the data} rather than an arbitary function computed on it. This, in turn, allows us to define the intrinsic entropy, or randomness, inherent in a single object expressed as a binary string, without any assumptions on the process that produced it. However, this notion of complexity does not in all respects match our intuitive idea of which objects are complex, and which objects are not.

@@

Over the years one idea in particular has been proposed in many different forms. We select a description for the data by two-part coding (describing a model first, and then the data in terms of the model), and we take the complexity of only the model part as a measure of the meaningful information. If we can find a single two-part description that we can objectively call the optimal description of the data, the complexity of the data part gives us an objective property of the data. While this idea has been proposed many times in many different guises, we will, for the purposes of this paper refer to it as \emph{sophistication}\footnotemark. The general problem of how to characterize the `richness' of the information in a string rather than the plain information content (to which sophistication is but one proposed solution), we will call the `problem of interesting information'.

\footnotetext{As the historical overview shows, sophistication is only one name for this idea, and not the first. The alternatives, however, are more awkward, and it has been used by different authors for different realizations of the same principle.}

The intuitive idea of sophistication makes one crucial demand: we can find an objectively `true' two-part description. In the framework of Komogorov complexity two observers cannot disagree by more than a constant about the information content of a dataset. In formal terms: the Kolmogorov complexity of a dataset $x$ is defined as the shortest effective description for $x$ in some Turing complete description language. This description language is usually expressed as a universal Turing machine $U(\bar\imath y) = \phi_i(y)$ where $\phi_i$ is the $i$'th partial computable function in some effective enumeration, and the $\bar\imath$ is a self-delimiting encoding of $i$. If two people disagree about the shortest effective description of $x$, we can interpret this as them using different description languages, ie. different universal Turing machines. The crucial aspect of Kolmogorov complexity, the starting theorem that makes it interesting as a function at all, is \emph{invariance}: two people using different description languages will get the same value for a string's Kolmogorov complexity, albeit up to a fixed constant. Without invariance, Kolmogorov complexity would be an arbitrary function. With invariance, we can be sure that we are talking about an intrinsic property of the data. The same consideration must be given to each proposal for formalising sophistication: is it invariant?

There are many approaches to sophistication, but the following pattern is common to all of them. We start with a space of \emph{models}. There can be finite sets, computable probability distributions or simple functions from strings to strings. The function perspective is the most generic, all other model classes can be translated to it. The approach to sophistication is to find a representation for a dataset $x$ in two parts: a computable function $f$ and and input $y$, such that $f(y) = x$. As with Kolmogorov complexity, the shorter a description is, the better. We take the shortest description and those close to it, and consider them \emph{candidates}. How close a description must be to the optimum differs between variants, it is often a pre-determined constant value. We then choose a single description from the candidate set and return its \emph{model complexity} as the sophistication. The method for determining model complexity varies. In almost all variants the candidate with the smallest model complexity determines the sophistication.

The implicit assumption behind sophistication is that this approach will separate the description of the string into a sturctural part, representing its regularities, and a stochastic part representing its non-deterministic properties. In other words, the string is cast as a random choice from a model for which it is `typical'.

@@

We first provide a chronological overview of the variants of sophistication. In the body of the paper, we will define each variant rigorously, and in non-chronological order.

The first suggestion that a representation of a string can be separated cleanly into a structural part, containing a description of its regularities, and a noise part, containing the residual information, came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{todo} and later more extensively discussed by Vit\'anyi \cite{todo}). Here, he proposed what has become known as the \emph{Kolmogorov structure function}. Briefly, this function operates on a space of two part representations for a given string $x$, consisting of some finite set $S$ (containing $x$) and the index of $x$ in $S$. The \emph{minimal sufficient statistic} is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity. \pb{TODO: lowest randomness deficienty.}

Koppel \cite{} built on this notion, coining the phrase \emph{sophistication}. He changed the class of `models' from finite sets to total functions and used Levin's monotone complexity, instead of prefix-free Kolmogorov complexity. This is likely to extend the theory to infinte strings, an approach which has been abandoned by subsequent authors.

While Koppel expands the model class from finite sets to functions, he maintains a restriction that the functions be total. This is a recurring theme: if the model class is expanded to all computable functions, the the universal Turing machine $U$ itself becomes a model. If we assume that $U$ computes the function $\phi_u$ in the standard enumeration of computable functions, and that $\bar\imath y$ is a shortest description for $x$ on $U$, then $\br{u}\br{\imath}y$ is also a description for $x$. This description has a constant model size ($|\br{u}|$), ie. every $x$ has constant sophistication. This `collapse' to a universal modle is the reason that most treatments of sophistication constrain the model class to exclude the universal Turing machine. This acts as a kind of force, pushing information into the model part of the two-part code. As we shall see, defining sophistication quickly becomes a balancing act, trying to match two forces which by themselves would push information one way or the other.

In 1996 Gell-Mann and Lloyd introduced `Effective complexity'\cite{gellmann1996information}. Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that firts the mold of sophistication. Effective complexity describes a binary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) and the entropy of the distribution. The reasoning is that, if the string is typical for the distribution, the entropy is sufficient to describe the string given the distribution, so that the length of the description of the distribution and the size of its entropy suffices as a two-part description. The two-part description which determines the effective complexity is chosen by placing a time bound (taken from the singleton description) on the computation of the string from the distribution. The hope is that this will move deep (ie. hard to compute) information to the model part, and shallow information to the data. In a later treatment\cite{gell2004nonextensive}, Gell-Mann and Lloyd select the decisive two-part representation from the candidates based on ``the average quantities judged to be important'', which seems to be a concious choice to make the measure subjective and context-dependant (sacrificing invariance).

G\'acs, Tromp and Vit\'anyi \cite{}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, as an attempt to found a statistical theory not on probabilistic notions, but on combinatorial ones. This echoes Kolmogorov's own hope for the theory of information:
\begin{quotation}
\noindent Information theory must precede probability theory and not be based on it. By the very essence of this discipline, the foundations of information theory must have a finite combinatorial character. \cite{kolmogorov1983combinatorial}
\end{quotation}
Algorithmic statistics studies binary strings and their possible two-part representations. In a later paper\cite{gacs2001algorithmic}, Vit\'anyi specifically studies the role of the minimal sufficient statistic as an expression of the `meaningful' information of a string (taking the name sophistication from Koppel). Like Koppel, Vit\'anyi restricts the model class to total functions, eliminating the universal model. He balances this force by allowing the total two-part description to be non-self delimiting, so that both the representation with all information in the model part, and the representation with everything in the residual part are less efficient than the best two-part coding.\sdr{This balancing act is not yet clear to the reader}

Antunes et al. \cite{} note that the constant cutoff used to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability. Small changes to this constant could shift the result sophistication by arbitrary amounts. They introduce a more elaborate criterion for the cutoff, calling this variant \emph{coarse sophistication}. As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth.

@@

\pb{Criticisms of sophistication}

@@

We now have the basic tools with which we set up our generic sophistication function. We first define the space of two-part representations for a given $x$, and the code length that each represenrtation achieves:

@@

\begin{definition}[Sophistication]
  The model complexity, meaningful information, effective complexity, facticity or \emph{sophistication},  of a string $x$ given a model class $\M$ and threshold function $\tau$ is
  \[
  \s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}\p
  \]\label{definition:sophistication}
\end{definition}

For a function like sophistication to be meaningful, it must satisfy certain properties. We will check all currently published variants against the following minimal requirements:

@@

For all variants but one, we have shown that all measures that are invariant are bounded. This leads us to the following conjecture. Let $\gamma(n)$ be a function representing the extent to which we care about differences between functions. That is, we will call two functions $f$ and $g$ $\gamma$-equal if $|f(x) - g(x)| \in O(\gamma(f(x)))$.

@@

For any sophistication function so constructed, and any granularity $\gamma$ we have that for any two acceptable numberings $g$ and $h$, either:
\begin{itemize}
  \item there exist $s_g$ and $s_h$ which are not are $\gamma$-equal (ie $s$ is not invariant),
  \item $s_g$ is $\gamma$-equal to $0$ or to the Kolmogorov complexity.
\end{itemize}
\end{conj}
In these terms, all instantiations of the Kolmogorov complexity, using different Universal Turing machines are $1$-equal. It may be unreasonable to require the sophistication to attain the same level of invariance (a log-invariant sophistication could still be useful). This conjecture allows such relaxations: if it can be shown to be true, sophistication must be considered fundamentally flawed. We hope that this conjecture will prompt authors to address the matter of invariance ciritcally, and to consider the possibility that sophistication and its ilk are an untenable proposition.

In the rest of this paper, we will analyse the existing variants one-by-one. We will conclude with the informal reasoning supporting this conjecture and consider its the consequences. What does this mean for the problem of meaningful information, what does it mean for algorithmic statistics and most importantly, what are the consequences for the idea of model selection in general?

@@

\subsection{Koppel Sophistication}

@@

This does not work because of the nickname problem! To fix this:

\begin{definition}
  The complexity of a partial recursive function is defined by
  \begin{align*}
    C(f) &:= \min\{C(i):\phi_i=f\}\\
    K(f) &:= \min\{K(i):\phi_i=f\}.
  \end{align*} \label{definition:model-complexity}
\end{definition}

These definitions make sense because:

\begin{lemma}[Invariance]
Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings. There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
\begin{align*}
K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
&= K^\psi(f).
\end{align*}
We can reverse $\phi$ and $\psi$ without loss of generality to achieve the same inequality the other way around, completing the proof.
\end{proof}

Thus, the complexity of a function depends on the choice of
enumeration by no more than a constant term. Note that this implies
that for any enumeration $\psi_1,\psi_2,\ldots$, the complexity $K(\psi_i)$
is less than the literal description length of the index
$|\bar\imath|$ up to a constant. For some numberings, there is a gap
between these two description methods; in those cases, the index is an
inefficient representation of the corresponding function. However it
is possible to construct numberings that are \emph{faithful} in the
sense that no such gap exists:

@@

\begin{itemize}
\item Model class is \emph{ensembles}. Equivalent to $\K$. Subsequent literature suggests that ensembles have finite support. 
\item Description method for models is $L(m) = K(m)$ (the minimization is not mentioned explicitly, but we'll give them the benefit of the doubt).
\item The threshold function is equal to the minimum two-part code: $\tau(x)=L^m(x)$.
\item Gell-mann and Lloyd acknowledge that representations exist with small and large models. The choice from the candidate set is made by applying a time-bound to the computation from the string to the model, but not to the decoding of the model itself. The idea is that this forces `deep' information into the model, and shallow information into the data. Presumably, the representation with the smallest model code under this constraint is chosen.   
\end{itemize} 
\pb{We can still manipulate the UTM so that all non-singleton representations are pushed out of the candidate set. Since the model side is allowed unbounded decoding time, this invalidates effective complexity.}

>>>>

\newcommand{\M}{\mathscr M}
\newcommand{\C}{\mathscr C}
\newcommand{\T}{\mathscr T}
\newcommand{\F}{\mathscr F}
\renewcommand{\P}{\mathscr P}
\newcommand{\K}{\mathscr K}
\newcommand{\X}{\mathscr X}
\newcommand{\B}{\mathbb B}
\newcommand{\D}{\Delta}
\newcommand{\N}{\mathbb N}
\newcommand{\tn}[1]{\textnormal{#1}}
\newcommand{\pair}[1]{\left\langle{#1}\right\rangle}
\newcommand{\concat}{\oplus}
\newcommand{\symb}[1]{\texttt{#1}}
\newcommand{\br}[1]{\overline{#1}}
\newcommand{\s}{\tn{soph}}
\newcommand{\num}[1]{#1_1, #1_2, \ldots}

@@

\begin{abstract}
Kolmogorov complexity is a sound expression of the information content of a binary string, but it clashes with our intuitive notions of information: the strings with maximal information content are not the ones we would regard as interesting, or meaningful. One attempt to capture this idea of `useful' information is \emph{sophistication}. It is a simple idea with many attempted formalizations: describe the string efficiently with a two-part coding approach, and take the complexity of the model as a measure of the amount of interesting information. Kolmogorov complexity is invariant to the ad-hoc choices in its construction: change them, and the result will not change more than a constant. We review the history of sophistication critically, asking whether invariance exists for the proposed sophistication functions as well. In most cases, we can show that it doesn't. We conjecture that no formalization of sophistication can be invariant, and we briefly review the consequences of this conjecture.
\end{abstract}

@@

\noindent Kolmogorov complexity is an objective measure of the information content of an object. Although its definition contains several arbitrary choices, these only affect its value up to a constant additive term. This property is crucial for the measure's success: it allows us to view Kolmogorov complexity as a \emph{property of the data} rather than an arbitrary function computed on it. This, in turn, allows us to define the intrinsic entropy, or randomness, inherent in a single object expressed as a binary string, without any assumptions on the process that produced it. However, this notion of complexity does not in all respects match our intuitive idea of which objects are complex, and which objects are not.

@@

Over the years one solution in particular has been proposed in many different forms. We select a description for the data by two-part coding; describing a model first, and then the data in terms of the model. The size in bits, used to describe the model is taken as a measure of the useful information in our string. While this idea has been proposed many times in many different guises, we will, for the purposes of this paper refer to it as \emph{sophistication}\footnotemark.

\footnotetext{As the historical overview shows, sophistication is only one name for this idea, and not the first. The alternatives, however, are more awkward, and sophistication has been used by different authors for different realizations of the same principle.}

The intuitive idea of sophistication makes one crucial demand: we can find an objectively value for the model complexity of a string, either by having a single objectively optimal two-part description, or a set of descriptions with equivalent model complexity. In the framework of Kolmogorov complexity, two observers cannot disagree by more than a constant about the information content of a dataset. In formal terms: the Kolmogorov complexity of a dataset $x$ is defined as the shortest effective description for $x$ in some Turing complete description language. This description language is usually expressed as a universal Turing machine $U(\bar\imath y) = \phi_i(y)$ where $\phi_i$ is the $i$'th partial computable function in some effective enumeration, and the $\bar\imath$ is a self-delimiting encoding of $i$. If two people disagree about the shortest effective description of $x$, we can interpret this as them using different description languages, ie. different universal Turing machines. The crucial aspect of Kolmogorov complexity, the starting theorem that makes it interesting as a function at all, is \emph{invariance}: two people using different description languages will get the same value for a string's Kolmogorov complexity, albeit up to a fixed constant. Without invariance, Kolmogorov complexity would be an arbitrary function. With invariance, we can be sure that we are talking about an intrinsic property of the data. The same consideration must be given to each proposal for formalising sophistication: is it invariant?

There are many approaches to sophistication, but the following pattern is common to all of them. We start with a space of \emph{models}. There can be finite sets, computable probability distributions or simple functions from strings to strings. The function perspective is the most generic, all other model classes can be translated to it. The approach to sophistication is to find a representation for a dataset $x$ in two parts: a computable function $f$ and and input $y$, such that $f(y) = x$. As with Kolmogorov complexity, the shorter a description is, the better. We take the shortest description and those close to it, and consider them \emph{candidates}. How close a description must be to the optimum differs between variants, it is often a pre-determined constant value. We then choose a single description from the candidate set and return its \emph{model complexity} as the sophistication. The method for determining model complexity varies. In almost all variants, the candidate with the smallest model complexity determines the sophistication.

The implicit assumption behind sophistication is that this approach will separate the description of the string into a structural part, representing its regularities, and a stochastic part representing its non-deterministic properties. In other words, the string is cast as a random choice from a model for which it is `typical'.

@@

We first provide a chronological overview of the existing variants of sophistication. In the body of the paper, we will define each variant rigorously, and in non-chronological order.

The first suggestion that a representation of a string can be separated cleanly into an encoding of its regular and its stochastic proeperties came from Kolmogorov himself, in a talk given at a conference in Tallinn in 1973 (first committed to paper by Cover \cite{cover1985kolmogorov} and later discussed more extensively discussed by Vereshchagin and Vit\'anyi \cite{vereshchagin2004kolmogorov}). Here, he proposed what has become known as the \emph{Kolmogorov structure function}. Briefly, this function operates on a space of two part representations for a given string $x$, consisting of some finite set $S$ (containing $x$) and the index of $x$ in $S$. The \emph{minimal sufficient statistic} is the set with the lowest complexity for which the length of the two-part representation reaches the prefix-free Kolmogorov complexity. It was first shown by Shen \cite{shen1983concept} that there are strings for which the best model cannot be small.

Koppel \cite{koppelSoph1988,koppel1991almost} built on this notion, coining the phrase \emph{sophistication}. He changed the class of `models' from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity. We expect that this change was made to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors. We will only consider the finite version (based on regular functions) in this paper.

While Koppel expands the model class from finite sets to functions,he does require that the functions be total. This is a recurring theme: if the model class is expanded to all computable functions, the universal Turing machine $U$ itself becomes a model. If we assume that $U$ computes the function $\phi_u$ in the standard enumeration of computable functions, and that $\bar\imath y$ is a shortest description for $x$ on $U$ (ie. $U(\bar\imath y) = x)$), then $\br{u}\br{\imath}y$ is also a description for $x$ on $\bar u$. This description has a constant model size ($|\br{u}|$), ie. every $x$ has constant sophistication. This `collapse' to a universal model is the reason that most treatments of sophistication constrain the model class to exclude the universal Turing machine. This makes it more expensive to move information out of the model. As we shall see, defining sophistication quickly becomes a balancing act, forcing information one way or the other.

In 1996 Gell-Mann and Lloyd introduced `Effective complexity' \cite{gellmann1996information}. Working from the perspective of statistical mechanics, they arrived (largely independently, it seems) at a notion that fits the mold of sophistication. Effective complexity describes a binary string in terms of a probability distribution over finite strings (an `ensemble' in the terminology of Gell-Mann and Lloyd) and the entropy of the distribution. The reasoning is that, if the string is typical for the distribution, the entropy is sufficient to describe the string given the distribution, so that the length of the description of the distribution and the size of its entropy suffices as a two-part description. The two-part description which determines the effective complexity is chosen by placing a time bound (taken from the singleton description) on the computation of the string from the distribution. The hope is that this will move deep (ie. hard to compute) information to the model part, and shallow information to the data. In a later treatment\cite{gell2004nonextensive}, Gell-Mann and Lloyd select the decisive two-part representation from the candidates based on ``the average quantities judged to be important'', which seems to be a conscious choice to make the measure subjective and context-dependant (sacrificing invariance).

G\'acs, Tromp and Vit\'anyi \cite{gacs2001algorithmic}, built on the idea of the structure function to establish a framework of \emph{algorithmic statistics}, as an attempt to found a statistical theory not on probabilistic notions, but on combinatorial ones. This echoes Kolmogorov's own hope for the theory of information:
\begin{quotation}
\noindent Information theory must precede probability theory and not be based on it. By the very essence of this discipline, the foundations of information theory must have a finite combinatorial character. \cite{kolmogorov1983combinatorial}
\end{quotation}
Algorithmic statistics studies binary strings and their possible two-part representations. In a later paper \cite{gacs2001algorithmic}, Vit\'anyi specifically studies the role of the minimal sufficient statistic as an expression of the `meaningful' information of a string (taking the name sophistication from Koppel). Like Koppel, Vit\'anyi restricts the model class to total functions, eliminating the universal model.

Antunes et al. \cite{antunes2009sophistication} note that the constant cutoff used to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability. Small changes to this constant could shift the result sophistication by arbitrary amounts. They introduce a more elaborate criterion for the cutoff, calling this variant \emph{coarse sophistication}. As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth.

@@

@@

We now have the basic tools with which we set up our generic sophistication function. We first define the space of two-part representations for a given string $x$, and the codelength that each representation achieves:

@@

\begin{definition}[Sophistication]
  The model complexity, meaningful information, effective complexity, facticity or \emph{sophistication},  of a string $x$ given a model class $\M$ and threshold function $\tau$ is
  \[
  \s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}\p
  \]\label{definition:sophistication}
\end{definition}
For a function like sophistication to be meaningful, it must satisfy certain properties. We will check all currently published variants against the following minimal requirements:

@@

For all variants but one, we show that all measures that are invariant are bounded. This leads us to the following conjecture. Let $\gamma(n)$ be a function representing a \emph{granularity}: the extent to which we care about differences between functions. That is, we will call two functions $f$ and $g$ $\gamma$-equal if $|f(x) - g(x)| \in O\left(\gamma\left(f(x)\right)\right)$.

@@

For any sophistication function so constructed, and any granularity $\gamma$ we have that for any two acceptable numberings $g$ and $h$, either:
\begin{itemize}
  \item there exist $\s_g$ and $\s_h$ which are not are $\gamma$-equal (ie $\s$ is not invariant),
  \item $s_g$ is $\gamma$-equal to $0$ or to the Kolmogorov complexity.
\end{itemize}
\end{conj}

In these terms, all instantiations of the Kolmogorov complexity, using different Universal Turing machines are $1$-equal. It may be unreasonable to require the sophistication to attain the same level of invariance (a log-invariant sophistication could still be useful). This conjecture allows such relaxations: if it is true, sophistication must be considered fundamentally flawed. We hope that this conjecture will prompt authors to address the matter of invariance critically, and to consider the possibility that sophistication and its ilk are an untenable proposition.

In the rest of this paper, we will analyse the existing variants one by one. We will conclude with the informal reasoning supporting this conjecture and consider its the consequences. What does this mean for the problem of meaningful information, what does it mean for algorithmic statistics and most importantly, what are the consequences for the idea of model selection in general?

@@

\subsection{Koppel's Sophistication}

@@

We start with sophistication as originally introduced by Koppel and Atlan \cite{koppelSoph1988,koppel1991almost}. Their sophistication is built on a two-argument universal Turing Machine $U(i, y) = \phi_p(y)$. They consider a pair $(i, y)$ a representation of a string $x$ if (a) $\phi_i$ is total, and (b) $\phi_i(y)$ has $x$ as a prefix.\footnotemark

\footnotetext{The original sophistication also allows infinite binary strings in the range of the computable functions considered. We ignore this aspect for the sake of simplicity, although our conclusions still hold with such functions included.}

This construction shows one of the first pitfalls of sophistication: using an inefficient representation for the model. We show that unless 
\begin{lemma}
Koppel and Atlan's sophistication is not both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
\end{lemma}
\begin{proof}
Assume that the sophistication is $\gamma$-invariant, for some $\gamma(n) \in o(n)$. We will show that this implies that it is bounded.

Let $\num{\phi}$ be the canonical acceptable numbering. Choose $\alpha < 1$ so that $\gamma(i) < \alpha i$ for sufficiently large $i$. Let $s: \N \to \B$ be a computable function such that $|s(i)| = \left\lceil(1-\alpha)^{-i}\right\rceil$.

We now define two new acceptable numberings $\num{\psi}$ and $\num{\xi}$:
\[
\psi_j(x) = 
\begin{cases}
	\phi_i(x) \;\text{if}\; j = s(2i) \\
	\infty\;\text{otherwise}
\end{cases}
\]
and 
\[
\xi_j(x) = 
\begin{cases}
	\phi_i(x) \;\text{if}\; j = s(2i+1) \\
	\infty \;\text{otherwise}\p
\end{cases}
\]
Let $\s_\psi$ and $\s_\xi$ be the sophistication functions built on these enumerations. Choose any $x$ and assume w.l.o.g. that $\s_\psi(x) < \s_\xi(x)$. By construction of the numberings $\psi$ and $\xi$, we have $\s_\psi(x)/ (1-\alpha)  \leq \s_\xi(x) + 1$, which gives us:
\begin{align*}
	\s_\xi(x) - \s_\psi(x) &\geq \alpha\, \s_\xi(x) - 1\\
	& \in \Omega(\s_\xi(x)) - 1\\
\end{align*}
while, by our original assumption $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, which contradicts the bound above.
\end{proof}

The main trick we use in this proof is that we can inflate the indexes arbitrarily. In short, the representation of the models is highly inefficient. For the definition of Kolmogorov complexity, this is not a problem, since we can switch to another universal machine at only a constant cost, but for the definition of sophistication, we must make sure that both parts of our two-part code are efficient.

We define the complexity of a given model as its shortest effective description: a description that can be used to compute the function for a given input:

\begin{definition}
  The complexity of a partial recursive function is defined by
  \begin{align*}
    C(f) &:= \min\{C(i):\phi_i=f\}\\
    K(f) &:= \min\{K(i):\phi_i=f\}.
  \end{align*} \label{definition:model-complexity}
\end{definition}
These definitions also appear in \cite{} and \cite{}, but are not discussed in depth. We will show that these properties are still invariant up to a constant (ie. complexity is a property of a function):
\begin{lemma}[Invariance]
Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings. There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
\begin{align*}
K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
&= K^\psi(f).
\end{align*}
We can reverse $\phi$ and $\psi$ without loss of generality to achieve the same inequality the other way around, completing the proof.
\end{proof}

This lemma shows that the complexity of a function depends on the choice of
enumeration by no more than a constant term. Note that this implies
that for any enumeration $\psi_1,\psi_2,\ldots$, the complexity $K(\psi_i)$
is less than the literal description length of the index
$|\bar\imath|$ up to a constant. For some numberings, there is a gap
between these two description methods; in those cases, the index is an
inefficient representation of the corresponding function. However it
is possible to construct numberings that are \emph{faithful} in the
sense that no such gap exists:

@@

\begin{itemize}
\item Model class is \emph{ensembles}. Equivalent to $\K$. Subsequent literature suggests that ensembles have finite support, which would mean that effective complexity reduces to the structure function.
\item Description method for models is $L(m) = K(m)$ (the minimization is not mentioned explicitly, but we'll give them the benefit of the doubt).
\item The threshold function is equal to the minimum two-part code: $\tau(x)=L^m(x)$.
\item Gell-mann and Lloyd acknowledge that representations exist with small and large models. The choice from the candidate set is made by applying a time-bound to the computation from the string to the model, but not to the decoding of the model itself. The idea is that this forces `deep' information into the model, and shallow information into the data. Presumably, the representation with the smallest model code under this constraint is chosen.   
\end{itemize}

\pb{\cite{ay2010effective} provides a more formal definition.}
\begin{theorem}
There exists a UTM so that for all $x$, $\s_\text{EC}(x) = K(x)$, up to a constant. 
\end{theorem}

Subsequent publications suggest that all ensemebke
